Designing decentralized topic-based pub/sub systems is a big research
challenge due to a number of desired system properties which are in
conflict with eachother. For example, making the
overlay robust is difficult without introducing too many redundant edges
in the network graph.  Many approaches to topic-based pub/sub have been
proposed the last
decade~\cites{Baehni:2004}{Castro:2002}{Chockler:2007}{Rahimian:2011}{Girdzijauskas:2010}{Matos:2010}{Wong:2008}{Zhuang:2001}.
And each have made trade-offs in an attempt to balance the system
properties against each other. In this chapter, we extend the
mini-survey found in~\cite{Setty:2012}, were we include a number of
additional protocols. Also, we will go more into detail regarding the
design challenges found in topic-based pub/sub system by illustrating
the charachteristics of these systems, as well as their shortcomings.

\section{Desired System Properties}

In order to provide correct and efficient delivery of notifications in a
decentralized OSN using topic-based pub/sub, a high number of system
properties are deemed desirable~\cite{Setty:2012}. More specifically,
these challenges include:

\begin{description}

    \item[Correct delivery]\hfill\\
        All notifications should be delivered to the
        correct recipient. Both false positives and false negatives should be
        avoided.

    \item[High hit-ratio during churn]\hfill\\
        Notifications are delivered to a very high
        percentage of subscribers in the presence of churn. In the
        absence of churn, notifications should be delivered to all
        subscribers. This is similar to correct delivery except for
        not taking false positives into account.

    \item[Fast recovery]\hfill\\
        The overlay should quickly recover from a
        period of churn. Nodes should be able to both leave and join the
        network gracefully, and nodes who are dead should be properly
        handled by the system.

    \item[Low average node degree]\hfill\\
        The overlay nodes should have a low
        node degree as possible to achieve scalability with regards to number of
        topics. The degree distribution should be as even as possible,
        in order to achieve load balancing.

    \item[Topic connectivity]\hfill\\
        The routing of an event only includes the
        subscribers who registered their interest for the topic. This is
        also known as \emph{relay-free routing}.

    \item[Scalability]\hfill\\
        The system should scale in terms of number of
        topics, number of nodes, number of topics a node is interested in and
        number of nodes interested in a topic.

    \item[Efficient dissemination]\hfill\\
        Event dissemination should have a low
        delay with little duplicate delivery, and the load of routing messages
        should be distributed fairly.

    \item[Low overlay maintenance cost]\hfill\\
        Managing the overlay topology
        should be as inexpensive as possible. Maintenance might include mending
        dissemination structures such as multicast trees when nodes fail, but
        also how to include joining nodes in the structure and allowing nodes to
        leave gracefully.

\end{description}

Designing a system with all of these properties
presents a challenge, as several of the desired characteristics are
fundamentally at odds with each other. Maintaining a low node degree
makes it difficult to maintain \emph{topic connectivity}, while
avoiding duplicate message delivery conflicts with being robust in the
presence of churn. There is also a trade-off in robustness and
reliability depending on the approach taken to disseminating messages.
Specialized overlays that build dissemination structures such as
multicast trees provide fast and reliable message delivery with no
duplication of messages. However, they are fragile and susceptible to
churn. Epidemic dissemination on the other hand is more robust, but does
not provide full reliability as it lacks deterministic delivery
guarantees.

There is also a trade-off between the navigability of the overlay and
the message overhead. Stanley Milgram famously demonstrated the
\emph{small-world phenomena} in~\cite{milgram1967small} where he showed
that any two participants in a network was likely to be connected
through a low number of intermediaries. Taking this phenomena into
consideration has proved to be a useful approach when constructing
decentralised overlays, as they provide a highly navigable network due
to the small average shortest path length. A popular approach is to
create one or more long jump links between nodes to provide better
routing capabilities. More specifically, these links are usually created
by utilizing a distance function in the name space, where the
probability of creating a link increases with the distance between them.
The subscription interest of such nodes are usually not taken into
consideration when creating such links. Consequently, the message
overhead in the system is increased as more relays are introduced in the
overlay.

Many existing systems suffer from shortcomings that originate from wanting
to include a high number of desired properties described above. This
motivates further research into these systems and how they compare in
terms of promoting these desirable properties.

\section{Handling Trade-offs}

Designers of existing approaches have been facing the challenges of
handling the trade-offs discussed in the previous section. One of these
challenges is building both a reliable and robust overlay. A naive
approach to this problem would be to create a separate overlay for each
topic as in TERA~\cite{Baldoni:2007}. However, this approach suffers from
poor scalability as the number of nodes and topics increase. Another
approach would be to structure the overlay by creating a spanning tree
per topic as seen in Scribe~\cite{Castro:2002}, Bayeux~\cite{Zhuang:2001}
and Magnet~\cite{Girdzijauskas:2010}. However, these structures are
conceptually fragile in the presence of churn, requiring mechanisms for
mending the structure when nodes fail. This increases the overhead of overlay
maintenance.  Also, the root node of the spanning tree represents a
single point of failure as well as a bottleneck in the system. This is
especially true for popular topics where all events must travel through
the root node. In Scribe, the root node is used as a \emph{rendezvous}
point for topics by using the routing capabilities of the underlying
Pastry DHT~\cite{Rowstron:2001}. Such dedicated nodes represent a
single point of failure in addition to being detrimental to load
balancing and scalability.

Minimizing the average node degree while simultaneously achieving a
\emph{topic-connected overlay} (TCO) is another difficult trade-off to
consider. Topic-connectivity is achieved when no other than the
nodes who registered their interest in a topic takes part in routing
events for that topic. The desired goal with regards to
topic-connectivity is not only to avoid routing events through
uninterested nodes, but also to minimize the node degree by reusing
links for several topics. This approach achieves better scalability with
regards to the number of topics in the system.  Also, it decreases the
message overhead incurred by both event dissemination and overlay
maintenance mechanisms such as heartbeat messages. In addition to this,
keeping an overlay topic-connected simplifies the message routing
mechanism as no designated relay or gateway node needs to be implemented
in the protocol such as in Scribe and Vitis.

ElastO~\cite{Chen:2013} propose an interesting approach to overlay
construction, where the goal is to construct a TCO while maintaining a low
node degree. The construction of the overlay is performed by a
centralized component which requires global knowledge, while the
maintenance of the overlay is performed in a distributed manner in
response to churn events. By using a centralized algorithm for overlay
construction, ElastO is able to provide a more optimal TCO than decentralized
solutions, while still maintaining the high performance of a
decentralized repair mechanism to handle node departure or arrival.

With regards to the reuse of links for several topics, an observed
correlation~\cite{Liu:2005} between subscription sets in practical
workloads is useful to consider when constructing overlays. This
observation is exploited in Poldercast in order to decrease the number of
links to maintain. Also, it is used as a basis for overlay construction
in both StaN~\cite{Matos:2010} and SpiderCast~\cite{Chockler:2007}.
However, these two protocols only provide a probabilistic guarantee that
the resulting overlay will be fully topic-connected. In contrast,
PolderCast claim deterministic guarantees of providing a TCO. However,
this relies on two factors: (1) that there is no churn in the system,
and (2) that the underlying Cyclon~\cite{Voulgaris:2005} protocol, which
is used for peer sampling in PolderCast, can guarantee a connected
overlay.  Consequently, the deterministic guarantees of PolderCast could
be questioned.\ daMulticast~\cite{Baehni:2004} on the other hand provide
a deterministic guarantee of topic-connectivity through quite a
different approach of overlay construction. More specifically,
daMulticast constructs a topic hierarchy, where events are disseminated
through gossiping each level of this hierarchy in a bottom-up approach.

As mentioned in section 2, several protocols attempts to create an overlay that
exhibits \emph{small-world properties}. In Vitis, the subcluster
together with the relay paths form an overlay similar to a small-world
networks, which decreases the routing delay, but includes
uninterested gateway and relay nodes. In Magnet, small-world properties
are provided by the underlying Oscar DHT~\cite{girdzijauskas2007oscar}
which cluster similar nodes together.

\subsection{Overlay construction}

There are several different approaches to overlay construction.
Structured approaches such as dissemination trees have already been
mentioned, but there are also unstructured approaches. In Quasar
\cite{Wong:2008}, a novel approach to event dissemination using random
walks removes the need for a structured overlay.  There are also hybrid
approaches to structuring overlays such as in ElastO~\cite{Chen:2013}
and Vitis~\cite{Rahimian:2011}. In ElastO, the bootstrapping of the
graph is performed by a centralized entity, while in Vitis nodes with
similar interests are clustered together. A topic in Vitis might consist
of several subclusters which are connected to each other through relay
paths. This creates an overlay that is similar to dissemination trees,
but where single nodes have been replaced with clusters of nodes.
However, the drawbacks are still similar to the ones found in systems
relying on multicast trees, as it relies on designated gateway nodes
within subclusters communicating with rendezvous nodes along the relay
path. In Poldercast~\cite{Setty:2012}, a structured ring per topic is
used in combination with a form of epidemic dissemination that resembles
gossiping. Publishers are themselves part of the ring of the topic they
publish, and the structures attempt to combined into themselves into a
single overlay through random links. Such an hybrid approach is an
attempt at balancing the reliability of a structured overlay with the
robustness of epidemic dissemination. When it comes to node degree
however, Poldercast might introduce hotspots in the system as the
distribution of random links might be skewed.

In Magnet, the aforementioned subscription correlation is used to build
dissemination trees such as the ones seen in Scribe and Bayeux. As
mentioned these structures are not ideal in dynamic systems as they
require maintenance. However, tree structures do have an advantage when
it comes communication overhead, as they in theory avoid any duplicate
messages. In practice however, duplicate messages might occur if a node
that is part of a tree is part of an underlying routing path to the root
node. For example, in Scribe, notifications are routed to the root node
of a topic tree using an underlying DHT. If a child node belonging to
this tree is part of the routing path, it will receive the same message
twice. Once while routing to the root, and once again from its parent
after the publication message has reached the root node. However, the
number of duplicates should still be lower than what is seen in systems
relying on epidemic dissemination schemes, such as daMulticast and
PolderCast. In such systems, the number of duplicate messages are indeed
higher, however, the increased number of control messages also has the
benefit of making these systems more resilient to churn. Furthermore,
there is usually an adjustable \emph{fanout parameter} in such systems
which can be manipulated in order to control the number of messages that
are forwarded by a node. Thus, there is some control over the amount of
communication overhead in these systems. Both PolderCast and daMulticast
include such a fanout parameter. Also, it bears mentioning that
structured overlays include communication overhead in the form of
structure maintenance and structure mending in the case of both node
failure as well as nodes leaving and joining the network. Control
messages such as heartbeats are commonplace in such protocols e.g.\ in
Scribe where the each non-leaf node in the multicast tree periodically
sends heartbeats to its children. This increases bandwidth consumption
and adds a higher communication overhead compared to unstructured
overlays where such maintenance usually is not required.

Node degree is another important issue to consider when designing
overlays. A low average node degree increases scalability as topics and
number of nodes in the system increases. In protocols who rely on an
underlying DHT, node degree is usually either constant or a logarithmic
function of the total number of nodes in the graph. Such is the case in
Bayeux which relies on the Tapestry DHT~\cite{tapestry}, or Scribe which
relies on the Pastry DTH~\cite{Rowstron:2001}. Other implementations
might have a constant node degree, which is the case in Vitis which
might result in the separation of a topic into subclusters. Some extreme
examples include TERA~\cite{Baldoni:2007} and
daMulticast~\cite{voulgaris:2007} which
has a node degree that grows in the order of the number of topics the
node has subscribed to.  In the worst case scenario, this is also the
case in PolderCast, as maintaining a low node degree depends on the
degree of correlation in the subscription sets.  Indeed, when using
workloads from Facebook, the node degree in PolderCast grows almost
linearly with subscription size, as shown in \cite{Setty:2012}. This
suggests a scalability issue in a scenario where the subscription
correlation is weak.

%{../tables/comp-overlay} <- vim gf
\input{tables/comp-overlay}

Table~\ref{table:comp-overlay} provides an overview over several
different state-of-the-art protocols, comparing their different system
properties such as topic connectivity and whether or not it takes
advantage of subscription correlation.  Note that
PolderCast has received the benefit of doubt in this table, and have
been marked as providing a deterministic guarantee of
topic-connectivity. Magnet has a different approach to the spanning tree
structures, where messages are disseminated bottom-up. This means that
the root node is not a rendezvous node according to  the traditional
definition~\cite{baldoni2005distributed}, but it is still conceptually
a single point of failure as it is responsible for propagating messages
back down the tree when it receives a message. In SpiderCast, there is a possibility
of the overlay forming into a pattern of highly connected clusters
inter-connected through a small number of links which we refer to as
\emph{weak bridges}. The node degree in SpiderCast also relies on the
\emph{K-coverage} parameter of the system, where, for each topic, a node
attempts to connect to $K$ neighbours who share the same interest.
Protocols who rely on a underlying DHT typically have a node degree
which grows logarithmically with the number of nodes in the system. The
exception is Magnet which leverages a DHT providing small-world
properties~\cite{girdzijauskas2007oscar}, creating a fixed node degree
that is independent of both subscription size and number of topics
\cite{Zhuang:2001}.  Note that the node degree in Quasar is omitted from
the table, as it is dependent on the implementation of the bloom filters
used to represent the neighbours. In ElastO, $\rho$ is a system parameter
which balances between average and maximum node degrees when choosing
new edges to recover from churn.

% Note that even though ElastO has a
% higher node degree than the centralized solutions, it provides a more
% optimal TCO.\

\section{Event dissemination}

In terms of event dissemination, it should be mentioned that some of
the systems described earlier do not concern themselves with this aspect, and
focus only on the construction and maintenance of the overlay
itself. More specifically, this includes SpiderCast, StaN and ElastO. Thus, any
discussion regarding dissemination technique or routing performance
will be irrelevant for these systems. For other systems however, a
comparison of techniques is in order.

As mentioned, the dissemination in systems relying on multicast trees
removes much of the message duplication and usually offers on average
dissemination of events in logarithmic time as is the case in Scribe,
Magnet and Bayeux. In Vitis, event dissemination is performed by
flooding inside the subcluster, while simultaneously forwarding the
event to other subclusters if needed. As mentioned, gossiping is the
main approach in both daMulticast and PolderCast.  Gossiping usually
implies an exponential dissemination speed, however, there might be
other implementation specific factors in play which inhibits this
property of gossiping. As an example, in PolderCast, skewed random link
distribution might be detrimental to the speed of the gossiping
protocol. The authors of Quasar~\cite{Wong:2008} propose quite a different approach
to event dissemination, as routing is performed by having nodes install
routing vectors in nearby overlay neighbours. Messages are disseminated
through random walks, which are directed towards the subscribers when
passing through a node with the relevant routing information. This
approach is likely to be highly robust against churn. However, as
observed in~\cite{Wong:2008} the hit ratio stagnates at 97\% in a
static system. This is due to a phenomenon where some group members
might be obscured by other members who absorb messages
from all directions. The authors suggest a solution by having node
periodically pull information from other nodes, but this introduces more
overhead in terms of network traffic and data processing.

%{../tables/comp-dissemination} <- vim gf
\input{tables/comp-dissemination}

Table~\ref{table:comp-dissemination} describes the routing properties of
the protocols discussed in this section. Note that protocols relying on
a DHT usually have an expected delay which is logarithmic or squared
logarithmic with the total number of nodes in the system. In Vitis, the
underlying DHT provide squared logarithmic routing complexity with the
total number of nodes divided by $k$, the number of long-range
neighbours. To the best of our knowledge, there is no evaluation of the
hit-ratio of Magnet or Bayeux, which is the reason of these being marked
as unknown. Message delay is defined as the expected path length of a
dissemination message in terms of number of hops. Systems relying on an
underlying DHT usually provides routing performance which is logarithmic
to the number of nodes in the system.  Magnet differs in its approach to
routing, as it relies on random walks with associated \emph{time-to-live} values.
However, as described in \cite{Girdzijauskas:2010} this value is usually
set to the logarithm of the total number of nodes in the system. The
average duplication factor describes the message overhead of the system,
where gossiping is usually dependent on a fanout system parameter. The
novel approach in Quasar means message overhead is dependent on the
number of parallel random walks initiated by a node. As described
earlier, protocols which create specialized dissemination structures, in
these cases spanning trees, have in theory no message duplication.

\section{Chapter Summary}

In this chapter, we extend the mini-survey found in~\cite{Setty:2012}.
We describe several different topic-based pub/sub protocols and their
properties. In particular, we focus on the several trade-offs that
researchers need to consider when designing such systems. We then
describe different system and how they relate to these trade-offs by
describing their properties and how it affects their performance in
terms of overlay construction and maintenance, as well as event dissemination.
